Species persistence under climate and fishing

### 2014-02-12 Writing/MS.tex
Working on Malin's comments, small edits

Knowing whether two disturbances are synergistic in their impacts for a given population is important for management. The co-occurrence of climate change-driven range shifts and fishing mean that there is the potential for synergistic interactions, which have been largely unexamined.  Previous work examining  how harvest and range shift interact has lacked the generality to examine how these two disturbances are likely to jointly affect population persistence. Many modeling studies have focused on how fish populations are likely to change in the future under a variety of fishing and climate scenarios, but the degree of detail and case-specificity makes it difficult to draw general conclusions. Here we have built a general model to examine how climate and harvesting interact to affect species persistence by incorporating dispersal and reproduction.

@@

We also examined frequently recommended management approaches: harvest thresholds and MPAs and whether these strategies would work to ensure species persistence. We found these two management strategies to both increase the population's biomass at equilibrium and improved its ability to persist. Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation \citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}. Our results, that spatial management increased the maximum harvesting rate at which the population could survive, support the idea that MPAs could be used to reduce the impact of harvesting. However we did not find any evidence that MPAs increased the climate velocity under which the population could persist. The second management approach we investigated, harvesting thresholds, are already widely implemented in fisheries management, and we found that this management tactic alleviates interactions between the two stressors.   While the management strategies only change harvesting practices and do not directly address the effects of climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement harvesting rules and place protected areas.

The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.  However, it  ignores many of the complexities present in marine fisheries. We do not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist over time. However, we found that qualitatively similar results about the interaction between climate and harvesting would hold for a model with a recruitment function with Allee effect. We also did not include age structure in our model. The effects of both harvesting and climate change may be different across different age classes; including this level of complexity is left for future work. Additionally, we did not include any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.  In addition to these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.  By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000}and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.

>>>>

Knowing whether two disturbances interact in their impacts for a given population is important for management. The co-occurrence of climate change-driven range shifts and fishing mean that there is the potential for synergistic interactions, which have been largely unexamined.  Previous work examining  how harvest and range shift interact has lacked the generality to examine how these two disturbances are likely to jointly affect population persistence. Many modeling studies have focused on how fish populations are likely to change in the future under a variety of fishing and climate scenarios, but the degree of detail and case-specificity makes it difficult to draw general conclusions. Here we have built a general model to examine how climate and harvesting interact to affect species persistence by incorporating dispersal and reproduction.

@@

We also examined whether frequently recommended management approaches ensure species persistence. We found increases in the population's biomass at equilibrium and an improved ability to persist. Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation \citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}. Our results, that spatial management increased the maximum harvesting rate at which the population could survive, support the idea that MPAs could be used to reduce the impact of harvesting. However we did not find any evidence that MPAs increased the climate velocity under which the population could persist. The second management approach we investigated, harvesting thresholds, are already widely implemented in fisheries management, and we found that this management tactic alleviates interactions between the two stressors.   While the management strategies only change harvesting practices and do not directly address the effects of climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement harvesting rules and place protected areas.

The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.  However, it  ignores many of the complexities present in marine fisheries. We do not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist over time. However, we found that qualitatively similar results about the interaction between climate and harvesting would hold for a model with a recruitment function with Allee effect. We also did not include age structure in our model. The effects of both harvesting and climate change may be different across different age classes; including this level of complexity is left for future work. Similarly, we did not include any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.  Besides these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.  By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000}and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.
